---
title: "Portfolio Project"
author: "Jaron Johnstone"
format: html
editor: source
---

## Section 1: Data Description

```{r}
#| label: load packages 
#| warning: FALSE
#| message: FALSE

library(ggplot2)
library(dplyr)
library(tidymodels)
library(glmnet)
library(tidyverse)
library(rsample)
library(naniar)
library(LearnPCA)
library(readr)
library(ExamPAData)
library(randomForest)
```

```{r}
#| label: import data
#| warning: FALSE
#| message: FALSE
auto_claim <- readr::read_csv(here::here("auto_claim.csv"))

### Imputing some missing data 
auto_claim <- auto_claim |>
  mutate(
    YOJ = ifelse(is.na(YOJ), median(YOJ, na.rm = TRUE), YOJ),
    INCOME = ifelse(is.na(INCOME), median(INCOME, na.rm = TRUE), INCOME),
    HOME_VAL = ifelse(is.na(HOME_VAL), median(HOME_VAL, na.rm = TRUE), HOME_VAL),
    SAMEHOME = ifelse(is.na(SAMEHOME), median(SAMEHOME, na.rm = TRUE), SAMEHOME)
  )
```

```{r}
#Convert Claim Flag to a factor variable 
auto_claim <- auto_claim %>%
  mutate(CLM_FLAG = as.factor(CLM_FLAG))

#Peak in to data 
head(auto_claim)
```

```{r}
#Stratified split to keep proportion of claim flags roughly equal between the two groups 
set.seed(839)
split <- initial_split(auto_claim, prop = 0.8, strata = CLM_FLAG)
train <- training(split)
test <- testing(split)
```

#### Description
The auto_claim data set comes from the SOA's Predictive Analytics Exam practice materials, it contains 10296 observations and 29 variables. This data was collected in order to promote predictive analytics education within an actuarial context. Each observation is an individual auto insurance policy. Some of the variables include demographic information of the policy holder such as age, gender, income, and education. Other variables include characteristics about the vehicle such as car type, blue book value, and car use. The response variable that is being investigated in this analysis will be the CLM_FLAG variable which indicates whether or not a policy has made a claim. About 27% of policies have a claim, while 73% do not. We will be exploring which policy holder and vehicle characteristics are associated with a higher likelihood of filing a claim.

Before splitting the data, missing values in INCOME, HOME_VAL, YOJ, and SAMEHOME were imputed using median imputation, preserving all 10296 observations. The data was then split in to a training set(80%, n = 8236) and a test set(20%, n = 2060). A stratified sample was used to ensure that the ratio of policies with claims vs policies with no claims were roughly equal between the two sets. The size of the each set is appropriate since there is ample data to train the models with while keeping enough data left over to test on.

## Section 2: Data Wrangling and Exploratory Data Analysis

The goal of this analysis is to identify which characteristics(either pertaining to the policyholder or their vehicle) are associated with a higher likelihood of making a claim. Other than a few missing data points, the data was largely clean and did not need any significant wrangling. This section will focus on the relationship between claim rate and key explanatory variables.

#### Univariate Summaries

```{r}
train |>
  select(AGE, INCOME, MVR_PTS, TRAVTIME, BLUEBOOK, HOME_VAL, KIDSDRIV) |>
  summary()
```

#### Univariate Plots

```{r} 
#| label: Unvariate Age 
#| 
ggplot(train, aes(x = AGE)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "white") +
  geom_vline(xintercept = 16, color = "red", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = 80, color = "red", linetype = "dashed", linewidth = 1) +
  labs(title = "Distribution of PolicyHolder Age", 
       x= "Age", y = "Count")
```

The distribution of policyholder age is roughly bell-shaped and centered around the mid 40s. The youngest policy holders are 16 years old and the oldest are 80 years old, reflecting a realistic spread of policyholder ages.

```{r}
#| label: Univariate MVR Points

ggplot(train, aes(x = MVR_PTS)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "white") +
  geom_vline(xintercept = 13, color = "red", linetype = "dashed", linewidth = 1) +
  labs(title = "Distribution of Motor Vehicle Record Points", 
       x= "MVR Points", y = "Count",
       caption = "Red dashed line indicates maximum value fo 13 points")

```

The distribution of vehicle record points is heavily right skewed with the majority of policyholders having 0 or 1 point, but a small number of drivers having as many as 13 points on their record(indicated by red line on the graph since it wasn't visible). This kind of distribution is typical with risk related variables where the majority of the data is low risk while a small amount of the data is high risk.

#### Bivariate Plots

```{r}
#| label: Bivariate Age vs Claim rate
train |>
  mutate(age_group = cut(AGE, 
                         breaks = c(16, 25, 35, 45, 55, 65, 80),
                         labels = c("16-25", "26-35", "36-45", 
                                    "46-55", "56-65", "66-80"),
                         include.lowest = TRUE)) |>
  filter(!is.na(age_group)) |>
  count(age_group, CLM_FLAG) |>
  group_by(age_group) |>
  mutate(pct = n / sum(n)) |>
  filter(CLM_FLAG == "Yes") |>
  ggplot(aes(x = age_group, y = pct)) +
  geom_col(fill = "steelblue") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Claim Rate by Age Group",
       x = "Age Group", y = "Claim Rate")
```

Claim rates are highest among the youngest drivers (16-25) at around 67%, and generally decline with age. This aligns with the well with the assumption that younger drivers are higher risk due to their lack of experience on the road . There is a slight uptick in the 56-65 group which may reflect increased reaction time issues associated with older age, though the rates fall again in the 66-80 age range, possibly due to retirees driving less frequently and thus have less exposure to accidents. 

```{r}
#| label: Bivariate Revoked Status vs Claim Rate
# Claim Rate by Revoked Status

train |>
  count(REVOLKED, CLM_FLAG) |>
  group_by(REVOLKED) |>
  mutate(percentage = n/sum(n)) |>
  filter(CLM_FLAG == "Yes") |>
  ggplot(aes(x = REVOLKED, y = percentage, fill == REVOLKED)) +
  geom_col(fill = "steelblue") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Claim Rate by License Revocation Status", 
       x = "License Revoked", y = "Claim Rate") +
  theme(legend.position = "none")
```

One of the strongest signals found in the data was the relationship between prior license revocation and claim rate. Drivers who have previously had their license revoked have a claim rate of 45%, which is nearly double the claim rate of 24% associated with drivers who have not had their license revoked before. This makes sense since a revoked license indicates serious driving violations, which is directly associated with higher accident risk.

```{r}
#| label: Bivariate Claim rate vs MVR points 
train |>
  mutate(mvr_group = cut(MVR_PTS,
                         breaks = c(-1, 0, 1, 3, 6, 13),
                         labels = c("0", "1", "2-3", "4-6", "7+"))) %>%
  count(mvr_group, CLM_FLAG) |>
  group_by(mvr_group) |>
  mutate(pct = n / sum(n)) |>
  filter(CLM_FLAG == "Yes") |>
  ggplot(aes(x = mvr_group, y = pct)) +
  geom_col(fill = "steelblue") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Claim Rate by MVR Points",
       x = "MVR Points", y = "Claim Rate")
```

When MVR points are grouped into ranges, a clear pattern emerges. Claim rates rise as points increase. Policyholders with zero points file claims at the lowest rate, while those with 7 or more points file claims at a dramatically higher rate. This makes intuitive sense since MVR points accumulate directly from traffic violations and at-fault accidents, making them one of the most direct measures of driving risk available to insurers.

```{r}
#| label: Bivariate Vehicle use vs. Claim rate
# Claim Rate by Vehicle Use

train |>
  count(CAR_USE, CLM_FLAG) |>
  group_by(CAR_USE) |>
  mutate(percentage = n/sum(n)) |>
  filter(CLM_FLAG == "Yes") |>
  ggplot(aes(x = CAR_USE, y = percentage, fill == CAR_USE)) +
  geom_col(fill = "steelblue") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Claim Rate by Car Use", 
       x = "Car Use", y = "Claim Rate") +
  theme(legend.position = "none")
```

Next we examine the claim rate of commercial verse private vehicles. Here we see that commercial vehicles have a claim rate of 35%, while privately owned vehicles have a claim rate of 22%. Commercial vehicles tend to be driven more frequently and often under rougher conditions, thereby having an increased exposure to accident risk when compared to privately owned vehicles.

```{r}
#| label: Bivariate Claim rate vs Educational Level 
#Claim rate By Education level

educ_levels <- c("<High School", "High School", "Bachelors", "Masters", "PhD")
train |>
  count(MAX_EDUC, CLM_FLAG) |>
  group_by(MAX_EDUC) |>
  mutate(percentage = n/sum(n)) |>
  filter(CLM_FLAG == "Yes") |>
  mutate(MAX_EDUC = factor(MAX_EDUC, levels = educ_levels)) |>
  ggplot(aes(x = MAX_EDUC, y = percentage, fill == MAX_EDUC)) +
  geom_col(fill = "steelblue") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Claim Rate by Education Level", 
       x = "Education level", y = "Claim Rate") +
  theme(legend.position = "none")
```

Examining maximum education levels of policy holders, with the exception of those who did not graduate high school, their is a clear inverse relationship between claim rate and education level. We see those with a high school education have a claim rate of 34% while those with a PhD have a claim rate of 15%. This pattern is probably due to other characteristics that are associated with achieving high levels of education, rather than the education itself.

```{r}
#| label: Bivariate Claim rate vs Car Type
# Claim Rate by Car Type

train |>
  count(CAR_TYPE, CLM_FLAG) |>
  group_by(CAR_TYPE) |>
  mutate(pct = n / sum(n),
         total = sum(n)) |>
  filter(CLM_FLAG == "Yes") |>
  ggplot(aes(x = reorder(CAR_TYPE, pct), y = pct)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = paste0("n = ", total), y = pct + 0.01), 
            vjust = 0, size = 3.5) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.45)) +
  labs(title = "Claim Rate by Car Type",
       x = "Car Type", y = "Claim Rate")
```

Finally, we examine the claim rate among different kinds of vehicles. Here, it is clear that sports cars and pickup trucks experience the highest claim rates at 34% and 32% respectively, while sedans have the lowest at 17%. The sedan and suv groups are the largest in the dataset (n = 2148 & n = 2309 respectively), making them the most statistically reliable estimates. Sports cars are likely associated with more aggressive/reckless driving and pickup trucks are more likely to be used as commercial vehicles, both of which are factors that are associated with an increased exposure to accident risk. On the other hand, sedans are often associated with more conservative everyday driving patterns, leading to lower accident rates.

## Section 3: Unsupervised Learning

Principal Component Analysis is a technique that aims to reduce the dimensionality of a large set of correlated variables in to a smaller set of uncorrelated variables called principal components. Each component is a linear combination of the original variables constructed in such a way that captures as much variation in the data as possible. The first component explains the most variation, the second component explains the second most variance, and so on. In this analysis, PCA was performed on 9 numeric variables from the set: AGE, INCOME, HOME_VAL, BLUEBOOK, TRAVTIME, MVR_PTS, NPOLICY, HOMEKIDS, and KIDSDRIV. All variables were standardized to ensure they contribute equally to the analysis.

```{r}
#Selecting numeric predictors only
pca_recipe <- recipe(
  ~ KIDSDRIV + TRAVTIME + BLUEBOOK + NPOLICY + MVR_PTS + AGE + HOMEKIDS + INCOME + HOME_VAL, 
  data = train) |>
  step_normalize(all_numeric_predictors()) |>
  step_pca(all_predictors())

pca_prep <- prep(pca_recipe)

tidy(pca_prep, number = 2)

tidy(pca_prep, number = 2, type = "variance")
```

```{r}
#| label: Scree Plot

tidy(pca_prep, number = 2, type = "variance") |>
  filter(terms == "percent variance") |>
  ggplot(aes(x = component, y = value)) +
  geom_col(fill = "steelblue") +
  labs(title = "Scree Plot - Variance Explained by each Principal Component",
       x = "Principal Component", y = "Precent Variance Explained")
```

The Scree plot shows the percentage of variance explained by each principal component. PC1 explains 23.8% of the variance and PC2 explains 16.2% with a noticeable drop off in the following components. In the end, we select 5 principal components, which account for about 73% of the variance in the model. This cut off balances dimensionality reduction (9 variables to 5) while also keeping useful information about the data.

```{r}
#| label: Loadings Plot

tidy(pca_prep, number = 2) |>
  filter(component %in% c("PC1", "PC2")) |>
  ggplot(aes(x = reorder(terms, value), y = value, fill = value > 0)) +
  geom_col()+
  facet_wrap(~ component) +
  coord_flip() +
  labs(title = "PCA Loadings for PC1 and PC2",
       x = "Variable", y = "Loading") +
  theme(legend.position = "none")
```

The loadings plot shows how much each variable contributes to each component. In PC1, income, home value, bluebook value, and age all load negatively while number of kids at home and kids driving load positively. This component captures the tension between wealth and family size. Wealthier, older policyholders fall on the negative side, while younger policyholders with more children fall on the positive side. Interestingly, age and the kids-related variables load in opposite directions, which may reflect a nonlinear relationship. It could be middle-aged policyholders may have the most children at home, while both very young and older policyholders tend to have fewer.
PC2 tells a different story: home value, income, and bluebook value load positively while age loads negatively. This component appears to capture younger, more affluent families, distinguishing them from older, wealthy policyholders who appear in PC1.

```{r}
#| label: Scores Plot

juice(pca_prep) |>
  bind_cols(train |> select(CLM_FLAG)) |>
  ggplot(aes(x = PC1, y = PC2, color = CLM_FLAG)) +
  geom_point(alpha = 0.3) +
  labs(title = "PC1 vs PC2 by Claim Status",
       x = "PC1(23.8%)", y = "PC2 (16.2%)", color = "Claim Filed")
```

Next is the scores plot that shows each policyholder on a plane defined by PC1 and PC2. Each point is colored based on whether or not that policyholder has made a claim. There is a lot of overlap between the two groups, which is to be expected since PCA is unsupervised and does not take claim status in to account in its formulation. There is a downward sloping, triangular shape rather than a symmetric cluster. This could be reflecting the fact that some variables are bounded at their maximum or minimum values, making certain combinations of PC1 and PC2 highly unlikely. Despite the overlap, there is a slight tendency for policies that have filed claims to appear on the positive side of PC1, consistent with the earlier assumption that younger/ lower income families tend to have higher claim rates.   

## Section 4: Supervised Learning

Two Classification models were used to predict whether a policyholder would file a claim. The first model was a logistic regression model with lasso regularization and the second model was a random forest. 5-fold cross validation was used to tune the hyperparameters.

```{r}
### convert character variables to factors 

factors <- c("CAR_USE", "CAR_TYPE", "RED_CAR", "REVOLKED", "GENDER", "MARRIED", "PARENT1", "JOBCLASS", "MAX_EDUC", "AREA")

train <- train |> mutate(across(all_of(factors), as.factor))
test <- test |> mutate(across(all_of(factors), as.factor))
```

```{r}
#Recipe
model_recipe <- recipe(CLM_FLAG ~ KIDSDRIV + TRAVTIME + BLUEBOOK + NPOLICY + MVR_PTS + AGE + HOMEKIDS + INCOME + HOME_VAL + YOJ + SAMEHOME + CAR_USE + CAR_TYPE + RED_CAR + REVOLKED + GENDER + MARRIED + PARENT1 + JOBCLASS + MAX_EDUC + AREA, data = train) |>
  step_dummy(all_nominal_predictors()) |>
  step_normalize(all_numeric_predictors())

#5 - fold Validation
set.seed(839)
cv_folds <- vfold_cv(train, v = 5, strata = CLM_FLAG)
```

#### Logistic Regression
Logistic regression is an extension of linear regression used to predict binary outcomes. In this case, it is whether a claim was filed or not. Regular linear regression predicts a continuous number, but we need a probability between 0 and 1. Logistic regression solves this by taking the linear combination of predictors and passing it through a sigmoid function, which constricts any value into a range between 0 and 1. Rather than a binary output of 0 or 1, the model outputs how likely it is that a claim was filed. If that predicted probability exceeds 50%, the policyholder is classified as Yes, otherwise No.

The model works by finding the best set of coefficients, one for each predictor, that maximizes the likelihood of correctly classifying the observed outcomes in the training data. Each coefficient tells us the direction and strength of a predictor's relationship with claim likelihood. For example, a positive coefficient on MVR points means that more points increases the predicted probability of a claim, while a negative coefficient on age means higher age decreases it. This makes logistic regression one of the most straightforward classification models available since it gives you a readable equation that explains exactly how each variable influences the prediction.

Lasso regularization was used to automatically penalize unimportant predictors in the model. This method simplifies the model while also preventing over fitting on the training data. The penalty in this case is referred to as 'lambda'. The higher lambda is, the higher the penalization is on the predictor variables, which results in a simpler model that may miss more patterns(more bias). A lower lambda fits the model more closely to the training data, but could result in over fitting(more variance). Higher variance means the model is more sensitive to changes in the training data, which could lead to model making poor predictions when presented with new data.The optimal lambda selected was 0.00234, this is a relatively small value which indicates that most of the predictors were useful.

```{r}
#logistic regression model 
log_reg_model <- logistic_reg(penalty = tune(), mixture = 1) |>
  set_engine("glmnet") |>
  set_mode("classification")

#work flow
log_wflow <- workflow() |>
  add_recipe(model_recipe) |>
  add_model(log_reg_model)

log_reg_grid <- grid_regular(penalty(), levels = 20)

set.seed(839)

log_reg_tune <- tune_grid(
  log_wflow,
  resamples = cv_folds,
  grid = log_reg_grid,
  metrics = metric_set(accuracy)
)

show_best(log_reg_tune, metric = "accuracy")
```

```{r}
#best penalty
best_penalty <- select_best(log_reg_tune, metric = "accuracy")

#finalzing workflow
log_reg_final <- log_wflow |>
  finalize_workflow(best_penalty)

#fit in training data
log_reg_fit <- log_reg_final |>
  fit(data = train)

#Predicting in test data
log_reg_results <- log_reg_fit |>
  predict(test) |>
  bind_cols(test |> select(CLM_FLAG)) |>
  metrics(truth = CLM_FLAG, estimate = .pred_class)

log_reg_results
```
The logistic regression achieved a mean cross validation accuracy of 78.5% (std error: 0.23%) and a test set accuracy of 79.1% 

#### Random Forest
In a random forest model, many decision trees are trained on a slightly different random sample (with replacement) of the data, which is a technique called bootstrapping. This method ensures each tree learns from a slightly different version of the data, making the trees diverse and less likely to all make the same mistakes. At every split in a tree, the model only considers a random subset of predictors rather than all of them, which is the key difference between a random forest and simply building many decision trees. Without this restriction, most trees would tend to use the same strong predictors at every split, making them too similar to each other and reducing the benefit of combining them. 

The number of predictors considered at each split is controlled by the hyperparameter mtry. Each tree then makes a prediction on a new data point by passing it through its series of yes/no questions, and the prediction that receives the most votes across all trees becomes the final prediction. This approach is powerful because even if individual trees make errors, those errors tend to cancel out when combined across hundreds of trees, resulting in a more stable and accurate model than any single tree could achieve on its own

```{r}
#Forest model
rf_model <- rand_forest(mtry = tune(), trees = 100) |>
  set_engine("randomForest") |>
  set_mode("classification")

#workflow
rf_wflow <- workflow() |>
  add_recipe(model_recipe) |>
  add_model(rf_model)

#tunring grid 
rf_grid <- grid_regular(mtry(range = c(2, 10)), levels = 3)

#Tuning 
set.seed(839)
rf_tune <- tune_grid(
  rf_wflow,
  resamples = cv_folds,
  grid = rf_grid
)
```

```{r}
show_best(rf_tune, metric = "accuracy")
```

```{r}
#best mtry
best_mtry <- select_best(rf_tune, metric = "accuracy")

#finalzing workflow
rf_final <- rf_wflow |>
  finalize_workflow(best_mtry)

#fit in training data
rf_fit <- rf_final |>
  fit(data = train)

#Predicting in test data
rf_results <- rf_fit |>
  predict(test) |>
  bind_cols(test |> select(CLM_FLAG)) |>
  metrics(truth = CLM_FLAG, estimate = .pred_class)

rf_results
```

The Random Forest achieved a mean cross validation accuracy of 78.3% (std error: 0.27%) and a test set accuracy of 78.1%

##### Model Performance and loss functions

```{r}
#Confusion matrix for logistic regression
log_reg_fit |>
  predict(test) |>
  bind_cols(test |> select(CLM_FLAG)) |>
  conf_mat(truth = CLM_FLAG, estimate = .pred_class)
```

```{r}
#Confusion matrix for random tree model
rf_fit |>
  predict(test) |>
  bind_cols(test %>% select(CLM_FLAG)) |>
  conf_mat(truth = CLM_FLAG, estimate = .pred_class)
```
Examining the confusion matrices reveals an important limitation of both models. The logistic regression correctly identified only 205 out of 548 actual claims in the test set, missing 343. Meaning the model failed to categorize 63% of the policies with claims as high risk. The random forest performed similarly, missing 344 out of 548 claims. Both models are conservative in predicting claims, likely because the training data has nearly three times as many non-claim policies as claim policies, so the models learn that predicting 'No' is usually correct.

From the perspective of an insurance company, this is a significant concern since each high risk observation that was classified as low risk represents a policyholder that was likely underpriced due to their underestimated risk exposure. While accuracy is roughly 79% between the two models, this is largely driven by correctly identifying the larger group of policyholders who have not made claims. A possible solution to this would be adjusting the classification threshold below 50%, resulting in more false positives in exchange for catching more real claims. 

#### Adjusting Threshold 
```{r}
# Get predicted probabilities for logistic regression
log_reg_probs <- log_reg_fit |>
  predict(test, type = "prob") |>
  bind_cols(test |> select(CLM_FLAG))

# Adjust threshold to 30%
log_reg_probs$.pred_class2 <- ifelse(
  log_reg_probs$.pred_Yes > 0.30,
  "Yes",
  "No"
) |>
  as.factor()

log_reg_probs |> conf_mat(
  truth = CLM_FLAG,
  estimate = .pred_class2
)
```

```{r}
# Get predicted probabilities for random forest
rf_probs <- rf_fit %>%
  predict(test, type = "prob") %>%
  bind_cols(test %>% select(CLM_FLAG))

# Adjust threshold to 30%
rf_probs$.pred_class2 <- ifelse(
  rf_probs$.pred_Yes > 0.30,
  "Yes",
  "No"
) |>
  as.factor()

rf_probs |> conf_mat(
  truth = CLM_FLAG,
  estimate = .pred_class2
)

```
To address the high false negative rate, the classification threshold was lowered from 50% to 30%. This means any policy with a predicted claim probability above 30% is flagged as high risk. With this threshold, the logistic regression correctly identified 392 out of 548 actual claims compared to only 205 at the default threshold, nearly doubling the number of claims caught. The random forest showed similar improvement, catching 393 claims. However this comes at the cost of an increased amount of false positives from 87 to 373 for logistic regression, meaning more low risk policyholders are incorrectly flagged as high risk and potentially overcharged. The optimal threshold ultimately depends on how the insurer weighs the cost of missing a claim against the cost of overcharging a low risk customer. 

## Conclusion and Insights
This analysis sets out to identify which policyholder and vehicle characteristics are most associated with an increased likelihood of filing an auto insurance claim. The exploratory data analysis revealed driving the history variables, like license revocation status and vehicle record points, were the strongest predictors of claim likelihood. These predictors are followed by vehicle characteristics such as car type and car use. Demographic factors such as age and education levels also showed meaningful patterns, with younger drivers and those with less education filing claims at higher rates. 

Both the logistic regression and random forest models achieved similar test set accuracies of approximately 79%, suggesting that the simpler logistic regression model was sufficient to capture the relationships in this dataset. However, accuracy alone is not the best measure of success in an insurance context. Correctly identifying high risk policyholders matters more than overall correctness, and both models struggled to catch the majority of actual claims at the default threshold. Therefore, lowering the classification threshold to 30%, reducing false negatives at the cost of increased false positives, is a logical step to take in this scenario.  

This analysis has several notable limitations. First, both models predict a binary outcome: whether a claim was filed or not. This is a simplification of how risk is treated in an insurance setting. In reality, policyholders vary not just in whether they file a claim but in how frequently they file claims and how severe those claims are. A binary classification model treats a policyholder who files one minor claim the same as one who files multiple large claims, missing important distinctions the risk levels of different policyholders. 

Several extensions of this analysis would be valuable in a real world actuarial context. Rather than predicting binary outcomes, models could predict claim frequency and claim severity seperately. This would give the insurer a better picture of the overall risk they are exposed to. Beyond predictions, models could be used to rank a policyholder's risk from low to high. This would aid underwriters in pricing new applicant's based on their risk profile. Such a system would allow insurers to more precisely match a policyholder's premium to their risk level, thus reducing the cross-subsubsidiaztion of premiums among policy holders. 

Overall, this analysis demonstrates that auto insurance claim likelihood can be predicted with decent accuracy using policyholder and vehicle characteristics alone. These findings serve as a strong foundation for building more sophisticated models that could price premiums accurately based on customer data.  
